# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

Loading data into R (file with the data must be in a current working directory):

```{r}
activity <- read.csv("activity.csv", header=T)
```

## What is mean total number of steps taken per day?

Histogram presenting the total number of steps taken each day.

```{r}
steps_per_day <- aggregate(steps ~ date, data = activity, sum)
```

```{r fig.width=7, fig.height=6}
barplot(steps_per_day$steps, names.arg = steps_per_day$date, xlab = "Date", ylab = "Number of steps per day", main="Total number of steps taken each day")
```

Mean of the total number of stpes taken per day amounts to:

```{r}
mean(steps_per_day$steps)
```

Median for the total number of stpes taken per day is:

```{r}
median(steps_per_day$steps)
```

## What is the average daily activity pattern?

```{r}
steps_interval <- aggregate(steps ~ interval, data = activity, mean)
```
``` {r fig.width=7, fig.height=6}
plot(steps_interval, type = "l", xlab="Interval", ylab="Average number of steps", main="Average number of steps taken across all days")
```

The maximum number of steps is taken in:
``` {r}
steps_interval$interval[which.max(steps_interval$steps)]
```
the 5-minute interval.

## Imputing missing values

Total number of rows with missing values amounts to:
```{r}
sum(is.na(activity))
```

Filling all of the missing values:
```{r}
steps_average <- aggregate(steps ~ interval, data = activity, mean)
impute_miss <- numeric()
for (i in 1:nrow(activity)) {
    case <- activity[i, ]
    if (is.na(case$steps)) {
        steps <- subset(steps_average, interval == case$interval)$steps
    } else {
        steps <- case$steps
    }
    impute_miss <- c(impute_miss, steps)
}
```

Creating new dataset with the missing values filled in:
```{r}
activity_filled <- activity
activity_filled$steps <- impute_miss
```

Histogram presenting the total number of steps taken each day after imputing missing values:
```{r}
steps_per_day_wo_NAs <- aggregate(steps ~ date, data = activity_filled, sum)
```
```{r fig.width=7, fig.height=6}
barplot(steps_per_day_wo_NAs$steps, names.arg = steps_per_day_wo_NAs$date, xlab = "Date", ylab = "Number of steps per day", main="Total number of steps taken each day after imputing NAs")
```

Mean of the total number of stpes taken per day amounts to:
```{r}
mean(steps_per_day_wo_NAs$steps)
```

Median for the total number of stpes taken per day is:
```{r}
median(steps_per_day_wo_NAs$steps)
```

## Are the diferences in activity patterns between weekdays and weekends?

Creating factor variable with two levels - "weekday" and "weekend":
```{r}
activity$date <- as.Date(activity$date, "%Y-%m-%d")
day <- weekdays(activity$date)
day_type <- vector()
for (i in 1:nrow(activity)) {
  if (day[i] == "Saturday") {
    day_type[i] <- "Weekend"
  } else if (day[i] == "Sunday") {
    day_type[i] <- "Weekend"
  } else {
    day_type[i] <- "Weekday"
  }
}
activity$day_type <- day_type
activity$day_type <- as.factor(activity$day_type)

steps_per_day_type <- aggregate(steps ~ interval + day_type, data = activity, mean)
names(steps_per_day_type) <- c("interval", "daylevel", "steps")

```

Creating appropriate plots:
```{r fig.width=7, fig.height=20}
library(lattice)
xyplot(steps ~ interval | day_type, steps_per_day_type, type = "l", layout = c(1, 2), 
       xlab = "Interval", ylab = "Number of steps taken per day")
```
